---
title: "Metagenomic ORF Detection and Annotation"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
options(scipen=999)
cpal=4

library(data.table,quietly = T)

#read in the ORF summary table
orf=fread(commandArgs(trailingOnly=T)[2],header = T,sep="\t")

#read in the gtf summary table
gtf=fread(commandArgs(trailingOnly=T)[1],header = T,sep="\t")

```

## ORF Detection

```{r, echo=FALSE, warning=FALSE}

noorfs=nrow(orf)
nocontigs=length(unique(orf$Contig))

```

A total of **`r noorfs`** ORFs with length **`r round(mean(orf$Size))` $\pm$ `r round(sd(orf$Size))`** (Mean $\pm$ SD)  were detected across **`r nocontigs`** contigs (across all samples).

There were **`r round(mean(table(orf$Contig)))` $\pm$ `r round(sd(table(orf$Contig)))`** (Mean $\pm$ SD) ORFs per contig (Range `r min(table(orf$Contig))` - `r max(table(orf$Contig))`).

```{r, echo=FALSE, warning=FALSE}

library(ggplot2,quietly = T)
library(RColorBrewer,quietly = T)
library(cowplot,quietly = T)
library(stringr,quietly = T)
theme_set(theme_cowplot())

#histogram of all ORF lengths

orflenhist = ggplot(orf,aes(Size))+geom_histogram(bins = 50,fill=brewer.pal(3,"Accent")[2])+theme_classic()+ylab("log10(ORF Count)")+xlab("ORF Size")+guides(fill=FALSE)+scale_y_log10()
plot(orflenhist)


```


Per sample stats are given below:

```{r, echo=FALSE, warning=FALSE}
orf$Sample=str_match(orf$Sample,"combined_annotations.dir/(\\S+).contigs.orf")[,2]

sample_count=table(orf$Sample)
sampc=data.frame(Names=names(sample_count),Counts=sample_count)
sampc=sampc[order(sampc$Counts.Freq),]

noorfspersamp=ggplot(sampc,aes(y=Counts.Freq,x=Names,fill=Names))+geom_bar(stat="identity")+coord_flip()+xlab("Sample")+ylab("No. ORFs")+guides(fill=FALSE)
plot(noorfspersamp)
orfsizepersamp=ggplot(orf,aes(x=Sample,y=log10(Size),fill=Sample))+geom_boxplot()+coord_flip()+ylab("log10(ORF Size)")+guides(fill=FALSE)
plot(orfsizepersamp)

orfsperconpersamp=table(orf$Contig,orf$Sample)

library(reshape2,quietly = T)
opcps=melt(orfsperconpersamp)
colnames(opcps)=c("Contig","Sample","No_ORFs")
opcpsplot=ggplot(opcps,aes(x=Sample,y=log10(No_ORFs),fill=Sample))+geom_boxplot()+coord_flip()+ylab("log10(No. ORFs per contig)")+guides(fill=FALSE)
plot(opcpsplot)

```

````{r, echo=FALSE}
taxfeats=c("kingdom","phylum","class","order","family","genus","species")
tax=gtf[which(gtf$feat_type%in%taxfeats),]

fun=gtf[-which(gtf$feat_type%in%taxfeats),]

```

## Taxonomic Annotation

```{r, echo=FALSE}
for(i in taxfeats){
  cat(paste0(length(unique(tax$feat_name[tax$feat_type==i]))," unique ",i," observed across all samples.\n"))  
}

sampcnt=rowSums(tax[,3:ncol(tax)])
sampcnt=data.frame(feat=tax$feat_name,Taxon_Lev=tax$feat_type,Count=sampcnt)
scplot=ggplot(sampcnt,aes(x=Taxon_Lev,y=Count,fill=Taxon_Lev))+geom_boxplot()+xlab("Taxonomic Level")+ylab("No. Samples Taxon Observed In")+guides(fill=FALSE)+coord_flip()
plot(scplot)

```

## Functional Annotation
 
```{r, echo=FALSE}

for(i in unique(fun$feat_type)){
  cat(paste0(length(unique(fun$feat_name[fun$feat_type==i]))," unique ",i," observed across all samples.\n"))  
}

fsampcnt=rowSums(fun[,3:ncol(fun)])
fsampcnt=data.frame(feat=fun$feat_name,Fun_Type=fun$feat_type,Count=fsampcnt)
fscplot=ggplot(fsampcnt,aes(x=Fun_Type,y=Count,fill=Fun_Type))+geom_boxplot()+xlab("Functional Annotation Type")+ylab("No. Samples Annotation Observed In")+guides(fill=FALSE)+coord_flip()
plot(fscplot)

```
