---
title: "Summary of Metagenomic Read Filtering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
options(scipen=999)
cpal=4
```

## Steps Run

```{r, echo=FALSE}
sfile=read.table(commandArgs(trailingOnly=T)[1],sep="\t",header=T)
rrna="was"
genome="was"
if(length(which(is.na(sfile$Post_rRNA_Filtering))) >0){rrna="was not"}
if(length(which(is.na(sfile$Post_Genome_Filtering))) >0){genome="was not"}
filecount=length(levels(sfile$File))
``` 

Ribosomal RNA filtering **`r rrna`** carried out and host genome filtering **`r genome`** carried out on **`r filecount`** files.

## Count Summaries

```{r, echo=FALSE}
if(rrna=="was not"){sfile$Post_rRNA_Filtering=sfile$Input}
if(genome=="was not"){sfile$Post_Genome_Filtering=sfile$Post_rRNA_Filtering}

om=mean(sfile$Input)
osd=sd(sfile$Input)
omin=min(sfile$Input)
omax=max(sfile$Input)

rm=mean(sfile$Post_rRNA_Filtering)
rsd=sd(sfile$Post_rRNA_Filtering)
rmin=min(sfile$Post_rRNA_Filtering)
rmax=max(sfile$Post_rRNA_Filtering)

gm=mean(sfile$Post_Genome_Filtering)
gsd=sd(sfile$Post_Genome_Filtering)
gmin=min(sfile$Post_Genome_Filtering)
gmax=max(sfile$Post_Genome_Filtering)

```

The **`r filecount`** input files contained **`r format(round(om),big.mark=",")` $\pm$ `r format(round(osd),big.mark=",")`** reads (Mean$\pm$SD), with a range of (**`r format(omin,big.mark=",")`-`r format(omax,big.mark=",")`**).

After rRNA filtering files contained **`r format(round(rm),big.mark=",")` $\pm$ `r format(round(rsd),big.mark=",")`** reads (Mean$\pm$SD), with a range of (**`r format(rmin,big.mark=",")`-`r format(rmax,big.mark=",")`**).

After host genome filtering files contained **`r format(round(gm),big.mark=",")` $\pm$ `r format(round(gsd),big.mark=",")`** reads (Mean$\pm$SD), with a range of (**`r format(gmin,big.mark=",")`-`r format(gmax,big.mark=",")`**).

```{r, echo=FALSE}
library(ggplot2)
library(reshape2)
colnames(sfile)=c("File","Input File","Post rRNA Filtering","Post Host Genome Filtering")
meltsfile=melt(sfile,id.vars =1)
readplot=ggplot(meltsfile,aes(x=variable,y=value,color=variable))+xlab("")+ylab("Read Count")+geom_boxplot(outlier.shape=NA)+geom_jitter()+theme_classic()+guides(fill=FALSE,color=FALSE)+scale_color_brewer(type="qual",palette = cpal)
readplot
```



## Relative Contribution of rRNA and Host Reads
```{r, echo=FALSE}
counts=matrix(nrow=nrow(sfile),ncol=3)
rownames(counts)=sfile$File
colnames(counts)=c("Unfiltered","rRNA","Host Genome")

counts[,3]=sfile$`Post rRNA Filtering`-sfile$`Post Host Genome Filtering`
counts[,2]=sfile$`Input File`-sfile$`Post rRNA Filtering`
counts[,1]=sfile$`Post Host Genome Filtering`
counts=data.frame(counts)
counts=counts[order(rowSums(counts)),]

percentages=data.frame((counts/rowSums(counts))*100)
percentages=percentages[order(percentages$Unfiltered),]

```

An average of **`r format(mean(percentages$Unfiltered),digits=3)`**% of reads remained after filtering, with an average of **`r format(mean(percentages$rRNA),digits=3)`**% and **`r format(mean(percentages$Host.Genome),digits=3)`**% aligning to rRNA and host genome sequences respectively.

```{r echo=FALSE}
percentmelt=melt(as.matrix(percentages))
countmelt=melt(as.matrix(counts))
newnames=c("Sample","Source","Value")
colnames(percentmelt)=newnames
colnames(countmelt)=newnames

percentmelt$Sample=factor(percentmelt$Sample,levels=rownames(percentages))
countmelt$Sample=factor(countmelt$Sample,levels=rownames(counts))

pplot=ggplot(percentmelt,aes(x=Sample,y=Value,fill=Source))+geom_bar(stat="identity")+coord_flip()+theme_classic()+
  xlab("Input File")+ylab("% Input Reads")+scale_fill_brewer(type="qual",palette=cpal)

cplot=ggplot(countmelt,aes(x=Sample,y=Value,fill=Source))+geom_bar(stat="identity")+coord_flip()+theme_classic()+
  xlab("Input File")+ylab("n Input Reads")+scale_fill_brewer(type="qual",palette=cpal)

pplot
cplot

```

## Output
Output files contain **`r format(round(gm),big.mark=",")` $\pm$ `r format(round(gsd),big.mark=",")`** reads (Mean$\pm$SD), with a range of (**`r format(gmin,big.mark=",")`-`r format(gmax,big.mark=",")`**).

```{r echo=FALSE}
counts=counts[order(counts$Unfiltered),]
countmelt$Sample=factor(countmelt$Sample,levels=rownames(counts))
outplot=ggplot(countmelt[which(countmelt$Source=="Unfiltered"),],aes(x=Sample,y=Value,fill=Source))+geom_bar(stat="identity")+theme_classic()+scale_fill_brewer(type="qual",palette=cpal)+guides(fill=FALSE)+coord_flip()+ylab("n Reads Passing All Filters")

outplot              
```



